%\documentclass[11pt,a4paper]{scrartcl}
%\documentclass[11pt,a4paper]{article}
\documentclass[11pt,letterpaper]{article}
\usepackage{tabularx}
\usepackage{array}
\usepackage{verbatim}
\newenvironment{sverbatim}{\begingroup\small\verbatim}{\endverbatim\endgroup}
\usepackage{lastpage}
\usepackage{fancyhdr}
\usepackage[nodayofweek]{datetime}
\input{mod_authnz_ibmdb2_info}
\usepackage[
  bookmarks=true,
  bookmarksnumbered=true,
  pdftitle={Documentation for mod authnz ibmdb2},
  pdfauthor={Helmut K. C. Tessarek},
  pdfcreator={\CommitHash{}},
  colorlinks=false,
  linkcolor={blue},
  citecolor={blue},
  backref=true,
  breaklinks=true
]{hyperref}

\newcommand{\stt}[1]{\tt \small #1}
\newcommand{\showtext}[2]{
\ifdata
#1#2
\fi
}

\begin{document}
\author{Helmut K. C. Tessarek}
\title{Documentation for\\
mod\_authnz\_ibmdb2\\
db2-hash-routines\\
scripts}
\date{\longdate{\formatdate{\CommitD{}}{\CommitM{}}{\CommitY{}}}}
\maketitle
\setcounter{page}{0}
\thispagestyle{empty}
\begin{abstract}
mod\_authnz\_ibmdb2 is an Apache authentication module using IBM\textsuperscript{\textregistered}{} DB2\textsuperscript{\textregistered}{} as the backend database for storing user and group information. The module supports several encryption methods.
\end{abstract}

\vfill

\begin{center}
\href{http://mod-auth-ibmdb2.sourceforge.net}{http://mod-auth-ibmdb2.sourceforge.net}\\

{\tiny \rm \showtext{Date: }\CommitDateTimeTz{} }
{\tiny \rm \showtext{Id: }\CommitHashAbbrev{} }

\end{center}
\newpage
\renewcommand{\thepage}{\roman{page}}
\tableofcontents
\newpage
\pagestyle{fancy}
\lhead{}
\chead{Documentation for mod\_authnz\_ibmdb2}
\rhead{}
\lfoot{{\tiny \rm \showtext{Date: }\CommitDate{}}}
\cfoot{\thepage\ / \pageref*{LastPage}}
\rfoot{{\tiny \rm \showtext{Id: }\CommitHashAbbrev{}}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\renewcommand{\thepage}{\arabic{page}}
\setcounter{page}{1}
\section{mod\_authnz\_ibmdb2}

\subsection{Building mod\_authnz\_ibmdb2}
Log in as root user.\\
\\
Change the {\tt DB2PATH} and the {\tt APXS} variables in the {\tt makemod} script according to your environment:
\begin{verbatim}
# Path settings

DB2PATH=/home/db2inst1/sqllib
APXS=/usr/local/apache/bin/apxs
\end{verbatim}
Set {\tt DB2PATH} to the directory where DB2 is accessed. This is usually the instance home directory.\\
\\
Set {\tt APXS} to the path that points to your apxs binary. The apxs binary is usually installed in the /$<$your apache home$>$/bin directory.\\
\\
After changing the above settings, run the script {\tt ./makemod}\\
\\
Last but not least the DB2 environment has to be set in the Apache startscript. This is done by sourcing the {\tt db2profile} script, which is located in {\tt DB2PATH}.
\subsection{Details on building mod\_authnz\_ibmdb2}
To build the module the Apache utility {\tt apxs} is used. The {\tt EXTRA\_LFLAG} needs to be specified so that the module will find the db2 library during runtime.
\begin{verbatim}
DB2PATH=/home/db2inst1/sqllib

EXTRA_LFLAG="-Wl,-G,-blibpath:$DB2PATH/lib"     (for AIX)
EXTRA_LFLAG="-Wl,-rpath,$DB2PATH/lib"           (for Linux)

apxs -c -ldb2 $EXTRA_LFLAG mod_authnz_ibmdb2.c
\end{verbatim}
If the {\tt sqlcli1.h} header file cannot be found, add the {\tt -I} option to specify the
directory where {\tt sqlcli1.h} can be found.
If the db2 library cannot be found, add the {\tt -L} option to specify the
directory where {\tt libdb2.so} can be found.\\

For example:
\begin{verbatim}
apxs -c -L/home/db2inst1/sqllib/lib        \
  -I/home/db2inst1/sqllib/include -ldb2    \
  $EXTRA_LFLAG mod_authnz_ibmdb2.c
\end{verbatim}
After creating the module, it has to be moved to the Apache module directory. This is also done with the apxs utility:

\begin{verbatim}
apxs -i mod_authnz_ibmdb2.la
\end{verbatim}
As the next step the DB2 environment has to be set in the Apache startscript. This is done by sourcing the {\tt db2profile} script, which is located in {\tt DB2PATH}.\\
\\
Finally, add the following directive to your httpd.conf and restart Apache:

\begin{verbatim}
LoadModule authnz_ibmdb2_module modules/mod_authnz_ibmdb2.so
\end{verbatim}

\subsection{Installing the Manpages}
There is a {\tt man} directory in the path, where you have extracted the mod\_authnz\_ibmdb2 package.\\
\\
Change into the man directory and run the script {\tt ./maninstall}
\newpage

\subsection{Description of the module}
\emph{mod\_authnz\_ibmdb2} is an Apache authentication module using IBM DB2 as the backend database for storing user and group information. The module is designed for Apache 2.2.x and later and is based on the new authentication/authorization framework.\\
\\
Here is a list of the new directives\footnote{see Appendix \ref{default}} that come with the module:\\
\\
\small
\begin{tabular}{@{} ll >{\raggedright\arraybackslash}p{ 50ex } @{}}
{\tt AuthIBMDB2Database} & & database name \\

{\tt AuthIBMDB2Hostname} & & database server hostname for uncataloged databases \\

{\tt AuthIBMDB2Portnumber} & & database instance port \mbox{(default: 50000)} \\

{\tt AuthIBMDB2User} & & user for connecting to the DB2 database \\

{\tt AuthIBMDB2Password} & & password for connecting to the DB2 database \\

{\tt AuthIBMDB2UserTable} & & name of the user table \\

{\tt AuthIBMDB2GroupTable} & & name of the group table \\

{\tt AuthIBMDB2NameField} & & name of the user column in the table \mbox{(default: {\tt username})} \\

{\tt AuthIBMDB2GroupField} & & name of the group column in the table \mbox{(default: {\tt groupname})} \\

{\tt AuthIBMDB2PasswordField} & & name of the password column in the table \mbox{(default: {\tt password})} \\

{\tt AuthIBMDB2CryptedPasswords} & & passwords are stored encrypted \mbox{(default: {\tt yes})} \\

{\tt AuthIBMDB2KeepAlive} & & connection kept open across requests \mbox{(default: {\tt yes})} \\

{\tt AuthIBMDB2Authoritative} & & lookup is authoritative \mbox{(default: {\tt yes})} \\

{\tt AuthIBMDB2NoPasswd} & & just check, if user is in usertable \mbox{(default: {\tt no})} \\

{\tt AuthIBMDB2UserCondition} & & restrict result set \\

{\tt AuthIBMDB2GroupCondition} & & restrict result set \\

{\tt AuthIBMDB2UserProc} & & stored procedure\footnotemark[2] for user authentication \\

{\tt AuthIBMDB2GroupProc} & & stored procedure\footnotemark[2] for group authentication \\

{\tt AuthIBMDB2Caching} & & user credentials are cached \mbox{(default: {\tt off})} \\

{\tt AuthIBMDB2GroupCaching} & & group information is cached \mbox{(default: {\tt off})} \\

{\tt AuthIBMDB2CacheFile} & & path to cache file \mbox{(default: {\tt /tmp/auth\_cred\_cache})} \\

{\tt AuthIBMDB2CacheLifetime} & & cache lifetime in seconds \mbox{(default: {\tt 300})} \\
\end{tabular}
\footnotetext[2]{see Appendix \ref{sp}}
\addtocounter{footnote}{1}
\normalsize

\newpage
\noindent
If {\tt AuthIBMDB2Authoritative} is {\tt Off}, then iff the user is not found in the database, let other authentication modules try to find the user. Default is {\tt On}. \\
\\
If {\tt AuthIBMDB2KeepAlive} is {\tt On}, then the server instance will keep the IBM DB2 server connection open.  In this case, the first time the connection is made, it will use the current set of Host, User, and Password settings.  Subsequent changes to these will not affect this server, so they should all be the same in every htaccess file.
If you need to access multiple IBM DB2 servers for this authorization scheme from the same web server, then keep this setting {\tt Off} -- this will open a new connection to the server every time it needs one.  The values of the database and various tables and fields are always used from the current {\tt .htaccess} file settings.\\
\\
If {\tt AuthIBMDB2NoPasswd} is {\tt On}, then any password the user enters will be accepted as long as the user exists in the database.\\ Setting this also overrides the setting for {\tt AuthIBMDB2PasswordField} to be the same as {\tt AuthIBMDB2NameField} (so that the SQL statements still work when there is no password at all in the database, and to remain backward-compatible with the default values for these fields.)\\
\\
For groups, we use the same {\tt AuthIBMDB2NameField} as above for the user ID, and {\tt AuthIBMDB2GroupField} to specify the group name.\\ {\tt AuthIBMDB2GroupTable} specifies the table to use to get the group info.  It defaults to the value of {\tt AuthIBMDB2UserTable}.  If you are not using groups, you do not need a {\tt groupname} field in your database, obviously.\\
\\
The optional directives {\tt AuthIBMDB2UserCondition} and \\{\tt AuthIBMDB2GroupCondition} can be used to restrict queries made against the User and Group tables. The value for each of these should be a string that you want added to the end of the where-clause when querying each table.
For example, if your user table has an {\tt active} integer field and you only want users to be able to login, if that field is 1, you could use a directive like this:\\
{\tt AuthIBMDB2UserCondition active=1}\\
\newpage
\noindent
If {\tt AuthIBMDB2UserProc} is set, the named \hyperlink{husersp}{stored procedure}\footnote{see Appendix \ref{usersp}} is responsible for returning the password of the user in question to the module. It must return exactly one value and row - the password. If set,\\ {\tt AuthIBMDB2UserTable}, {\tt AuthIBMDB2NameField}, {\tt AuthIBMDB2PasswordField}, {\tt AuthIBMDB2UserCondition} are ignored. If {\tt AuthIBMDB2NoPasswd} is {\tt On}, then the username has to be returned instead of the password. The stored procedure must have the following parameter format:
\begin{verbatim}
CREATE PROCEDURE user_procedure_name ( IN VARCHAR, OUT VARCHAR )
\end{verbatim}
If {\tt AuthIBMDB2GroupProc} is set, the named \hyperlink{hgroupsp}{stored procedure}\footnote{see Appendix \ref{groupsp}} is responsible for returning the groups the user in question belongs to. It must return an open cursor to the result set. If set,\\ {\tt AuthIBMDB2GroupTable}, {\tt AuthIBMDB2NameField}, {\tt AuthIBMDB2GroupField}, {\tt AuthIBMDB2GroupCondition} are ignored. The stored procedure must have the following parameter format:
\begin{verbatim}
CREATE PROCEDURE group_procedure_name ( IN VARCHAR )
\end{verbatim}
If {\tt AuthIBMDB2Caching} ist set to {\tt On}, the user credentials are cached in a file defined in {\tt AuthIBMDB2CacheFile} and expires after {\tt AuthIBMDB2CacheLifetime} seconds.\\
\\
If {\tt AuthIBMDB2GroupCaching} ist set to {\tt On}, the group information is cached in a cache file that is named like the file specified in {\tt AuthIBMDB2CacheFile} but with the extension {\tt .grp}. The cache expires after {\tt AuthIBMDB2CacheLifetime} seconds.
\newpage
\subsection{Examples}
First create the two tables within DB2:
\begin{verbatim}
CREATE TABLE WEB.USERS (
    USERNAME VARCHAR(40) NOT NULL,
    PASSWORD VARCHAR(40) );

ALTER TABLE WEB.USERS
    ADD PRIMARY KEY (USERNAME);

CREATE TABLE WEB.GROUPS (
    USERNAME  VARCHAR(40) NOT NULL,
    GROUPNAME VARCHAR(40) NOT NULL );

ALTER TABLE WEB.GROUPS
    ADD PRIMARY KEY (USERNAME, GROUPNAME);
\end{verbatim}
Then you will have to insert records into the two tables:
\marginpar{\footnotesize \flushleft {\tt bcrypt} is a User Defined Function that is explained in the \mbox{db2-hash-routines} part of this documentation.}
\begin{verbatim}
INSERT INTO WEB.USERS (username, password)
    VALUES ('test', bcrypt('testpwd'));
INSERT INTO WEB.GROUPS (username, groupname)
    VALUES ('test', 'admin');
\end{verbatim}
Then add the following lines to your httpd.conf:
\begin{verbatim}
<Directory "/var/www/my_test_dir">
    AuthName                    "DB2 Authentication"
    AuthType                    Basic
    AuthBasicProvider           ibmdb2

    AuthIBMDB2User              db2inst1
    AuthIBMDB2Password          ibmdb2
    AuthIBMDB2Database          auth
    AuthIBMDB2UserTable         web.users
    AuthIBMDB2NameField         username
    AuthIBMDB2PasswordField     passwd

    AuthIBMDB2CryptedPasswords  On
    AuthIBMDB2KeepAlive         On
    AuthIBMDB2Authoritative     On
    AuthIBMDB2NoPasswd          Off

    AuthIBMDB2GroupTable        web.groups
    AuthIBMDB2GroupField        groupname

    require                     group admin
    AllowOverride               None
</Directory>
\end{verbatim}
If you want to use \hyperlink{hsps}{stored procedures} and caching, the directives would look like this:
\begin{verbatim}
<Directory "/var/www/my_test_dir">
    AuthName                     "DB2 Authentication"
    AuthType                     Basic
    AuthBasicProvider            ibmdb2

    AuthIBMDB2User               db2inst1
    AuthIBMDB2Password           ibmdb2
    AuthIBMDB2Database           auth
    AuthIBMDB2UserProc           user_sp
    AuthIBMDB2GroupProc          group_sp

    AuthIBMDB2Caching            On
    AuthIBMDB2GroupCaching       On

    require                      group admin
    AllowOverride                None
</Directory>
\end{verbatim}
\newpage

% Include section: db2-hash-routines
\input{routines}
\newpage

% Include section: scripts
\input{scripts}
\newpage

\section{GIT access}
The git repositories can be cloned from github with the following instruction set:
\begin{verbatim}
git clone https://github.com/tessus/mod_authnz_ibmdb2.git
git clone https://github.com/tessus/mod_auth_ibmdb2.git
git clone https://github.com/tessus/db2-hash-routines.git
\end{verbatim}
You can also browse the repositories via the web:\\
\\
\begin{tabular}{@{} lr @{}}
mod\_authnz\_ibmdb2 & \url{https://github.com/tessus/mod_authnz_ibmdb2}\\ [0.5ex]
mod\_auth\_ibmdb2 & \url{https://github.com/tessus/mod_auth_ibmdb2}\\ [0.5ex]
db2-hash-routines & \url{https://github.com/tessus/db2-hash-routines}\\ [0.5ex]
\end{tabular}
\newpage

% Include section: FAQ
\input{faq}
\newpage

% Include section: Links
\input{links}
\newpage

\begin{appendix}
\section{directives and default values} \label{default}
\vspace{5ex}
\begin{tabular}{@{} lll @{}}
directive & & default value \\
\hline
& & \\
{\tt AuthIBMDB2Database} & & -- \\[0.5ex]
{\tt AuthIBMDB2Hostname} & & -- \\[0.5ex]
{\tt AuthIBMDB2Portnumber} & & {\tt 50000} \\[0.5ex]
{\tt AuthIBMDB2User} & & -- \\[0.5ex]
{\tt AuthIBMDB2Password} & & -- \\[0.5ex]
{\tt AuthIBMDB2UserTable} & & -- \\[0.5ex]
{\tt AuthIBMDB2GroupTable} & & -- \\[0.5ex]
{\tt AuthIBMDB2NameField} & & {\tt username} \\[0.5ex]
{\tt AuthIBMDB2GroupField} & & {\tt groupname} \\[0.5ex]
{\tt AuthIBMDB2PasswordField} & & {\tt password} \\[0.5ex]
{\tt AuthIBMDB2CryptedPasswords} & & {\tt yes} \\[0.5ex]
{\tt AuthIBMDB2KeepAlive} & & {\tt yes} \\[0.5ex]
{\tt AuthIBMDB2Authoritative} & & {\tt yes} \\[0.5ex]
{\tt AuthIBMDB2NoPasswd} & & {\tt no} \\[0.5ex]
{\tt AuthIBMDB2UserCondition} & & -- \\[0.5ex]
{\tt AuthIBMDB2GroupCondition} & & -- \\[0.5ex]
{\tt AuthIBMDB2UserProc} & & -- \\[0.5ex]
{\tt AuthIBMDB2GroupProc} & & -- \\[0.5ex]
{\tt AuthIBMDB2Caching} & & {\tt off} \\[0.5ex]
{\tt AuthIBMDB2GroupCaching} & & {\tt off} \\[0.5ex]
{\tt AuthIBMDB2CacheFile} & & {\tt /tmp/auth\_cred\_cache} \\[0.5ex]
{\tt AuthIBMDB2CacheLifetime} & & {\tt 300} \\[0.5ex]
\end{tabular}
\newpage

% Include reference: UDF and SP Reference
\input{udf_sp_ref}
\newpage

\hypertarget{hsps}{}
\section{Stored Procedure Support} \label{sp}
Stored procedures can minimize the network traffic and with regard to the authentication module configuration they can ease the administration. The module supports two types of stored procedures: one for user authentication and one for group authentication.\\
\\
For the following 2 sections we use these 3 tables:
\\
\begin{verbatim}
CREATE TABLE WEB.USERS (
    USERNAME VARCHAR(40) NOT NULL,
    PASSWORD VARCHAR(40) );

ALTER TABLE WEB.USERS
    ADD PRIMARY KEY (USERNAME);

CREATE TABLE WEB.GROUPS (
    GROUPNAME VARCHAR(40) NOT NULL,
    ACTIVE    INTEGER     NOT NULL );

ALTER TABLE WEB.GROUPS
    ADD PRIMARY KEY (GROUPNAME);

CREATE TABLE WEB.MAPPING (
    USERNAME  VARCHAR(40) NOT NULL,
    GROUPNAME VARCHAR(40) NOT NULL );

ALTER TABLE WEB.MAPPING
    ADD PRIMARY KEY (USERNAME, GROUPNAME)
    ADD FOREIGN KEY (USERNAME) REFERENCES WEB.USERS (USERNAME)
    ADD FOREIGN KEY (GROUPNAME) REFERENCES WEB.GROUPS (GROUPNAME);
\end{verbatim}
\newpage
\hypertarget{husersp}{}
\subsection{user authentication} \label{usersp}
The stored procedure for user authentication is responsible for returning the password of the user in question to the module. It must return exact one value - the password. If {\tt AuthIBMDB2NoPasswd} is {\tt On}, then the username has to be returned instead of the password.\\
\\
The stored procedure must have the following parameter format:
\begin{verbatim}
CREATE PROCEDURE user_procedure_name ( IN VARCHAR, OUT VARCHAR )
\end{verbatim}
Example:
\begin{verbatim}
CREATE PROCEDURE user_sp
(IN v_username VARCHAR(40), OUT v_password VARCHAR(40))
LANGUAGE SQL
BEGIN
  SELECT password INTO v_password FROM web.users
  WHERE username = v_username;
END@
\end{verbatim}
If {\tt AuthIBMDB2NoPasswd} is {\tt On}, then the stored procedure would have to look like this:
\begin{verbatim}
CREATE PROCEDURE user_sp
(IN v_username VARCHAR(40), OUT v_password VARCHAR(40))
LANGUAGE SQL
BEGIN
  SELECT username INTO v_password FROM web.users
  WHERE username = v_username;
END@
\end{verbatim}
\newpage
\hypertarget{hgroupsp}{}
\subsection{group authentication} \label{groupsp}
The stored procedure for group authentication is responsible for returning the groups the user in question belongs to. It must return an open cursor to the result set.\\
\\
The stored procedure must have the following parameter format:
\begin{verbatim}
CREATE PROCEDURE group_procedure_name ( IN VARCHAR )
\end{verbatim}
Example
\begin{verbatim}
CREATE PROCEDURE group_sp
(IN v_username VARCHAR(40))
LANGUAGE SQL
DYNAMIC RESULT SETS 1
BEGIN
  DECLARE res CURSOR WITH RETURN FOR
  SELECT m.groupname FROM web.groups g, web.mapping m
  WHERE m.groupname = g.groupname AND
        m.username = v_username AND
        g.active = 1;

  OPEN res;
END@
\end{verbatim}

\end{appendix}

\begin{center}
\vspace{\fill}\ \newline
{\tiny \rm \showtext{Date: }\CommitDateTimeTz{} }
{\tiny \rm \showtext{Id: }\CommitHash{} }
\end{center}

\end{document}
